---
title: "blognotes1"
author: "Carter Young"
format: html
editor: visual
---

### Notes 28/01/2024: Date formats in R
https://www.r-bloggers.com/2013/08/date-formats-in-r/

#### if dates stored as characters:

use **as.Date** & specify format of date
    ex. %m/%d/%y

seems to be part of base R--are there any packages that might do it differently?

trying it out:
```{r}
dates <- c("28/01/24")
betterDates <- as.Date (dates, format = "%d/%m/%y")
betterDates

```
looks like it worked!

#### if dates are numerical & am importing thru excel:
"*For Excel on Windows, the origin date is December 30, 1899 for dates after 1900. (Excel’s designer thought 1900 was a leap year, but it was not.) For Excel on Mac, the origin date is January 1, 1904.*" now that's weird but ok

need to define the origin date... so origin = 1904-01-01? 

*from Mac Excel: dates <- c(29367, 37078) betterDates <- as.Date(dates,origin = "1904-01-01")*

can format other date formulas, like days of the week
```{r}
dates <- c("Mon Jan 29")
format(betterDates,
  "%a %b %d")

```
I'm not sure why it gave me today's date. I'll look into that

there's a way to make sure two-digit years are the correct century, otherwise R will make assumptions about what's 1900s vs 2000s

note to self-figure out what the keybind is to select all for staging. not sure why it isn't working on this mac